Finally Tagless, Partially Evaluated

### 2008-11-14 tagless/tagless-final.tex
Go over letter to the editor

The abstract type constructors in Figure~\ref{fig:inductive} exemplify
the amount of polymorphism that our technique requires of the
metalanguage in order to represent a given object language.  Generally,
our technique represents term constructions (such as~$+$) by applying
abstract functions (such as \texttt{add}) and represents type
constructions (such as~$\to$) and typing judgments (namely `$:$') by
applying abstract type constructors (such as \texttt{darr} and
\texttt{repr}).  Therefore, it requires the metalanguage to support
enough polymorphism to abstract over the interpretation of each
inference rule for well-typed terms and well-kinded types.  For example,
to encode System~F's term generalization rule
\begin{equation*}
    \begin{prooftree}
        \[ [\alpha:\star] \proofoverdots e : t \]
        \justifies \Lambda\alpha.\,e : \forall\alpha.\,t \text,
    \end{prooftree}
\end{equation*}
the metalanguage must let terms (representing $\Lambda\alpha.\,e$)
abstract over terms (interpreting~$\Lambda$) that are polymorphic both
over type constructors of kind $\star\to\star$ (representing $t$ with
$\alpha$ free) and over polymorphic terms of type
$\forall\alpha\colon\star\dotso$ (representing $e$ with $\alpha$ free).
These uses of
higher-rank and higher-kind
polymorphism let us type-check and compile object terms separately from
interpreters.  This observation is consistent with the role of
polymorphism in the separate compilation of modules
\citep{shao-typed}.

>>>>

The abstract type constructors in Figure~\ref{fig:inductive} exemplify
the amount of polymorphism that our technique requires of the
metalanguage in order to represent a given object language.  Generally,
our technique represents term constructions (such as~$+$) by applying
abstract functions (such as \texttt{add}) and represents type
constructions (such as~$\to$) and typing judgments (namely `$:$') by
applying abstract type constructors (such as \texttt{darr} and
\texttt{repr}).  Therefore, it requires the metalanguage to support
enough polymorphism to abstract over the interpretation of each
inference rule for well-typed terms and well-kinded types.  For example,
to encode System~F's term generalization rule
\begin{equation*}
    \begin{prooftree}
        \[ [\alpha:\star] \proofoverdots e : t \]
        \justifies \Lambda\alpha.\,e : \forall\alpha.\,t \text,
    \end{prooftree}
\end{equation*}
the metalanguage must let terms (representing $\Lambda\alpha.\,e$)
abstract over terms (interpreting~$\Lambda$) that are polymorphic both
over type constructors of kind $\star\to\star$ (representing $t$ with
$\alpha$ free) and over polymorphic terms of type
$\forall\alpha\mathord:\mathord\star\dotso$ (representing $e$ with $\alpha$ free).
These uses of
higher-rank and higher-kind
polymorphism let us type-check and compile object terms separately from
interpreters.  This observation is consistent with the role of
polymorphism in the separate compilation of modules
\citep{shao-typed}.
